---
title: "Tidy Tuesday - Cocoa Beans"
author: "John Oliver"
date: '2022-01-21'
slug: tidy-tuesday-01-21-22
categories: null
tags: null
---

```{r, echo = F, include = F, warning = F}
# load packages
library(tidytuesdayR)
library(rworldmap)
library(dplyr)
library(ggplot2)
library(rnaturalearth)
library(rnaturalearthdata)
library(rgeos)
library(RColorBrewer)
```


This week marks the first week I will be participating in Tidy Tuesday. Tidy Tuesday, for those unfamiliar, is a weekly project hosted at https://github.com/rfordatascience/tidytuesday. Each Monday, a new data set is posted, and users around the world are challenged to dive in and find interesting information. There is no structure to it. You are simply handed a data set and told to get after it.

The project leader will give descriptions of variables, so you're not totally in the dark. Participants are encouraged to share their findings on Github, or on Twitter with the hashtag #TidyTuesday. It's an awesome way to establish a community of people who are trying to learn.

This weeks data set had everything to do with cocoa beans and chocolate. Specifically, this data looked into how users rated specific chocolate bars. Within the provided data set, each observation referred to one chocolate candy bar manufacturer. Recorded among each manufacturer were things such as location of manufacturing plant, average rating of their bar, and source location of where they were receiving their cocoa beans from. With no sense of direction and a loaded dataset, I was ready to dive into my first #TidyTuesday challenge.

# Producers

The process of constructing the world's perfect chocolate bar starts with cocoa beans. The manufactures turn these beans into the end product - a delicious candy bar. It's easy to praise the manufacturer when looking at the end product. In my eyes, however, the real heroes are the people growing these beans. Without them, this process would never occur!

That being said, I decided to look into where exactly these cocoa beans were being grown. To be honest, when I started this analysis, I had no idea where this stuff occurs. I assumed South America? To address my ignorance, I decide to produce a map that would tell me which countries grew beans and which did not. 

```{r, echo = F, include = F, warning = F}
# load data
choc_df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv')

# world dataset for plotting in ggplot
world_df <- rnaturalearth::ne_countries(scale = "medium",
                                        returnclass = "sf")

# mutate country names from chocolate dataset to match that of rnatural earth df
# remove county "Blend" as I can not figure out which country this is referring to
countries_fixed = choc_df %>% 
  mutate(updated_countries = recode(choc_df$country_of_bean_origin,
                                    "Burma" = "Myanmar",
                                    "Congo" = "Democratic Republic of the Congo",
                                    "DR Congo" = "Democratic Republic of the Congo", 
                                    "Martinique" = "Central African Republic" ,
                                    "Principe" = "Central African Republic",
                                    "Sao Tome" = "Central African Republic",
                                    "Sao Tome & Principe" = "Central African Republic" ,
                                    "St. Lucia"   = "Saint Lucia",
                                    "St.Vincent-Grenadines"  ="Saint Vincent and the Grenadines",
                                    "Sulawesi" = "Indonesia",
                                    "Sumatra"  ="Indonesia",
                                    "Tanzania" ="United Republic of Tanzania",
                                    "Tobago"  = "Trinidad and Tobago" ,
                                    "Trinidad"   = "Trinidad and Tobago" ,
                                    "U.S.A." = "United States of America")) %>% 
  filter(updated_countries != "Blend") 

# give a value of 1 to all countries that produce cocoa beens
world_choc <- world_df %>% 
  mutate(produces_choc = ifelse(admin %in% countries_fixed$updated_countries, "Produces",
                                "Does not produce"))
```

```{r, echo = F}
ggplot() +
  geom_sf(data = world_choc, aes(fill = produces_choc)) +
  scale_fill_manual(name = "Produces cocoa beans",
                    values = c("white", "red")) + 
  theme(panel.background = element_rect(fill = "aliceblue"))
```

